An Invariant Framework for Conducting Reproducible Computational Science

### 2015-01-09 intro.tex
Some rewording

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIS become an authoritative, encapsulated and executable records of computations, especially computations whose results are destined for publication and/or re-use, and this can be shared easily. 
With this approach, oftentimes, the resulting image is too big in size for share-ability purposes. An alternative light-weight form of virtualization allows encapsulating just the application software, along with all its necessary dependencies into a self contained package, thus separating , and not including, the underlying operating system on which it is executed. This encapsulation is achieved by interposing application system calls, and copying necessary dependencies (data, libraries, code) into a layered filesystem. While both forms of virtualization provides an excellent mechanism for sandboxing the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result.

Since reproducibility is significantly about comprehensive documentation, virtualization approaches in its current form only make it easy to capture the computations, but not preserve the computation so that it is easy to understand, install, or alter the implicit dependencies that are part of computation, especially as dependencies and external software components evolve or become deprecated.
Software delivery mechanisms, such as Linux Containers and Docker have recently emerged that address several of the reproducible challenges that arise due to use of virtualization approaches.

In this paper, we examine how software delivery mechanisms can be combined with traditional virtualization approaches for reproducibility, especially light-weight virtualization approaches. To conduct a thorough examination, we consider a real-world complex high energy physics application, independently developed by a group, but must be reproduced  so that the entire high energy physics community can benefit from the analysis obtained from the application. 
We describe challenges in reproducing this application, and consider the extent to which reproducibility can be satisfied with light-weight virtualization approaches and software delivery mechanisms. We then propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms. The framework uses light-weight virtualization for capturing a computational environment, but preserves the environment using a software delivery approach, so it can be re-executed later. By using standardized delivery mechanisms and expanding the application for future uses, the framework makes it invariant to temporal software changes that may occur in near future. We measure the performance overhead of combining light-weight virtualization with software delivery approaches. The combination imposes additional overhead due to double virtualization, but saves on context switches, since every system call is not interposed.

>>>>

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated and executable records of computations, especially computations whose results are destined for publication and/or re-use, and the VMIs can be shared easily \cite{}. 
With this approach, oftentimes, the resulting image is too big in size for share-ability purposes. An alternative light-weight form of virtualization allows encapsulation of the application software, along with all its necessary dependencies into a self-contained package, but does not make it executable.  It thus separates, and does not include the underlying operating system on which the package is executed. The encapsulation of the self-contained package is achieved by interposing application system calls, and only copying the necessary dependencies (data, libraries, code) into the package, which also makes it light-weight than a VMI. While both forms of virtualization provide mechanisms for sandboxing the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result.

Since reproducibility is significantly about comprehensive documentation, virtualization approaches in their current form only make it easy to capture the computations. Preserving the computations, so that they are easy to understand, install, or alter implicit dependencies that are part of computation, especially as dependencies and software components evolve or become deprecated, is not effectively addressed. There are two approaches to address the preservation challenge. By either introducing tools that help document dependencies and provide software attribution within VMIs or packages, or alternatively by using software delivery mechanisms, such as, centralized package management, linux containers and Docker framework. We examined the first approach previously in \cite{}. In this paper we examine the the second approach, that is, examine how the documentation and preservation standards that are part of software delivery mechanisms can be combined with traditional virtualization approaches for reproducibility. In particular we consider the light-weight virtualization approaches, because, we believe that the combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers. 
% they may help the This  We that must also include effective  to deliver the software to a wide audience. We examine the fire
%have recently emerged  several of the reproducible challenges that arise due to use of virtualization approaches.

To conduct a thorough examination, we consider a real-world complex high energy physics application, independently developed by a group, but must be reproduced  so that the entire high energy physics community can benefit from the analysis obtained from the application. The collaborative nature of in which high energy physics science is conducted is a 
We describe challenges in reproducing this application, and consider the extent to which reproducibility can be satisfied, independently with light-weight virtualization approaches and software delivery mechanisms. We then propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms. The framework uses light-weight virtualization for capturing a computational environment, but preserves the environment using a software delivery approach, so it can be re-executed later. By using standardized delivery mechanisms and expanding the application for future uses, the framework makes it invariant to temporal software changes that may occur in near future. We measure the performance overhead of combining light-weight virtualization with software delivery approaches. The combination imposes additional overhead due to double virtualization, but saves on context switches, since every system call is not interposed.
